The Ranked Nearest Neighbour Interchange space

### 2019-05-23 paper.tex
update proof convexity - there's still a bit missing

@@

\begin{algorithm}[H]
\caption{FIND\_PATH($T_1,T_2$)}
\label{alg:find_path}
\begin{algorithmic}[1]
	\STATE $T := T_1$
	\STATE $p := (T_1)$
	\FOR {$r = 1, \dots, n-2$}
		\STATE Let $S$ be the cluster associated with the node $v$ in $T_2$ with $rank(v) = r$
		\WHILE {$rank(MRCA_T(S))>r$}
			\STATE Update $T$: Decrease $rank(MRCA_T(S))$ by an $\rnni$ move \label{alg:line:move_set_down}
			\STATE $p = p+T$
		\ENDWHILE
	\ENDFOR
	\RETURN p
\end{algorithmic}
\end{algorithm}

@@

\begin{lemma}
	There are two caterpillar trees in $\rnniu$ with distance $\frac{1}{2}n^2-\frac{3}{2}n+1$.
	\label{conj:caterpillar_diameter}
\end{lemma}

@@

    In general, deleting taxon $n+1$ and its parent in every tree on any path $p$ from $T$ to $T'$ results in a path $p'$ from $T|_n$ to $T'|n$.
    Such a path $p'$ from $T|_n$ to $T'|n$ does not contain the $n-1$ $\rnni$ moves for moving the parent of $n+1$ down from the root to the cherry, so it is at least $n-1$ $\rnni$ moves shorter than $p$.
    If we assume that there is a path from $T$ to $T'$ of length less than $\Delta(\rnniu_{n+1}) = \frac{n(n-1)}{2}$, we could delete taxon $n+1$ and the $n-1$ moves of it's parent and receive a path from $T|_n$ to $T'|_n$ of length less than $\frac{n(n-1)}{2} - (n-1)$ = $\frac{(n-1)(n-2)}{2}$.
    So it would be $d(T|_n, T'|_n) < \Delta(\rnniu_n)$.
    Since this is a contradiction to $d(T|_n, T'|_n) = \Delta(\rnniu_n)$, it must be $d(T,T') = \Delta(\rnniu_{n+1}) = \frac{n(n-1)}{2} $.

@@

\begin{proof}
    We prove the Theorem by induction on the length $l$ of a shortest caterpillar path between two caterpillar trees $T_1$ and $T_2$.

    Induction basis: $l = \Delta(\rnniu) = \frac{(n-1)(n-2)}{2}$\\
    With Lemma~\ref{lemma:caterpillar_dist=diameter} one can see that for each pair of caterpillar trees $T_1$, $T_2$ with distance $d(T_1,T_2) = \Delta(\rnniu)$ there is a caterpillar path from $T_1$ to $T_2$ between these trees.

    Induction step: $l+1 \to l$\\
    Induction hypothesis: For each pait of caterpillar trees with distance $l+1$ there is a shortest path between these trees that only consists of caterpillar trees.
    Let $T_1$ and $T_2$ be caterpillar trees with caterpillar distance $l$ \todo{define 'caterpillar distance'}.
    If $l \neq \Delta(\rnniu)$, there must be a pair of taxa $x,y$ that are neighboured in $T_2$ and have the same order in $T_1$ and $T_2$, which means that $rank(p_x) > rank(p_y)$ or $rank(p_x) < rank(p_y)$ holds in both trees $T_1$ and $T_2$.
    Then it follows that the tree $T_2'$ that equals $T_2$, but with exchanged taxa $x$ and $y$, has distance $l+1$ to $T_1$. \todo{really??}
    So there must be a shortest path from $T_2'$ to $T_1$ that only consists of caterpillar trees.
    Since it does not matter in which order the pairs of taxa that need to exchange their position do exchange on a path between two caterpillar trees, $T_2$ is on a shortest caterpillar path from $T_2'$ to $T_1$.
    Therefore, it must be $d(T_1, T_2) = l$.

>>>>

%caterpillar path, shortest caterpillar path, caterpillar distance (def currently in proof of lemma:caterpillar_dist=diameter)

@@

\begin{algorithm}[H]
\caption{FIND\_PATH($T_1,T_2$)}
\label{alg:find_path}
\begin{algorithmic}[1]
	\STATE $T := T_1$
	\STATE $p := (T_1)$
	\FOR {$r = 1, \dots, n-2$}
		\STATE Let $S$ be the cluster associated with the node $v$ in $T_2$ with $rank(v) = r$
		\WHILE {$rank(MRCA_T(S))>r$}
			\STATE Update $T$: Decrease $rank(MRCA_T(S))$ by an $\rnni$ move \label{alg:line:move_set_down}
			\STATE $p = p+T$
		\ENDWHILE between these trees
	\ENDFOR
	\RETURN p
\end{algorithmic}
\end{algorithm}

@@

\begin{lemma} between these trees
	There are two caterpillar trees in $\rnniu$ with distance $\frac{1}{2}n^2-\frac{3}{2}n+1$.
	\label{conj:caterpillar_diameter}
\end{lemma}

@@

    In general, deleting taxon $n+1$ and its parent in every tree on any path $p$ from $T$ to $T'$ results in a path $p'$ from $T|_n$ to $T'|n$.
    Such a path $p'$ from $T|_n$ to $T'|n$ does not contain the $n-1$ $\rnni$ moves for moving the parent of $n+1$ down from the root to the cherry, so it is at least $n-1$ $\rnni$ moves shorter than $p$.
    If we assume that there is a path from $T$ to $T'$ of length less than $\Delta(\rnniu_{n+1}) = \frac{n(n-1)}{2}$, we could delete between these trees taxon $n+1$ and the $n-1$ moves of it's parent and receive a path from $T|_n$ to $T'|_n$ of length less than $\frac{n(n-1)}{2} - (n-1)$ = $\frac{(n-1)(n-2)}{2}$.
    So it would be $d(T|_n, T'|_n) < \Delta(\rnniu_n)$.
    Since this is a contradiction to $d(T|_n, T'|_n) = \Delta(\rnniu_n)$, it must be $d(T,T') = \Delta(\rnniu_{n+1}) = \frac{n(n-1)}{2} $.

@@

\begin{proof}
    We prove the Theorem by backwards induction on the length $l$ of a shortest caterpillar path between two caterpillar trees $T$ and $T'$.

    Induction basis: $l = \Delta(\rnniu) = \frac{(n-1)(n-2)}{2}$\\
    With Lemma~\ref{lemma:caterpillar_dist=diameter} one can see that for each pair of caterpillar trees $T$, $T'$ with distance $d(T,T') = \Delta(\rnniu)$ there is a caterpillar path that is a shortest path from $T$ to $T'$.

    Induction step: $l+1 \to l$\\
    Induction hypothesis: For each pair of caterpillar trees with distance $l+1$ there is a shortest path between these trees that only consists of caterpillar trees.
    Let $T$ and $T'$ be caterpillar trees with caterpillar distance $l$.
    Since $l < \Delta(\rnniu) = \frac{(n-1)(n-2)}{2}$, there is one loop $k$ in the Bubble Sort algorithm from $T'$ to $T$ where there are less than $n-2-k$ swaps of neighboured taxa.
    This means that the taxon $a$ that has parent with rank $n-2-k$ in $T'$ does not need to move all the way up starting at the cherry.
    Hence, there is a taxon $b$ that has parent with rank one less than the parent of taxon $a$ in $T'$ that does not exchange with $a$ on the path to $p$. \todo{$a$ and $b$ do not need to be neighbours in $T'$!}
    Let $T''$ be the caterpillar tree resulting from $T'$ by exchanging $a$ and $b$.
    It follows that $T'$ is on a shortest caterpillar path from $T''$ to $T$ and therefore, $d_c(T,T'') = d_c(T,T') + 1 = l+1$.
    We can apply the induction hypothesis on $T$ and $T''$ and know that $d(T,T'') = l+1$.
    If it was $d(T,T') < l$, it would hold $d(T,T'') \leq d(T,T') + d(T',T'') < l + 1$ which contradicts the induction hypothesis.
    We can conclude that $d(T,T') = l$.
